---
title: "Lab 04 - Data Visualization"
output:
  html_document:
    html_preview: false
link-citations: yes
---

```{r setup, message=FALSE, echo = FALSE, warning=FALSE, eval = TRUE}
library(data.table)
library(leaflet)
library(tidyverse)
```

# Learning Goals

-   Read in and prepare the meteorological dataset
-   Create several graphs with different `geoms()` in `ggplot2`
-   Create a facet graph
-   Conduct some customizations of the graphs
-   Create a more detailed map using `leaflet()`

# Lab Description

We will again work with the meteorological data presented in lecture.

**The objective of the lab is to examine the association between weekly average dew point temperature and wind speed in four regions of the US and by elevation.**

# Steps

### 1. Read in the data

First download and then read in with data.table:fread()

This is the same met_data as we used in lab 03, so you can choose to move that file to your current project directory instead of re-downloading.

```{r checking-file, include=FALSE}
fn <- "https://raw.githubusercontent.com/JSC370/jsc370-2022/main/labs/lab03/met_all.gz"
if (!file.exists("met_all.gz"))
  download.file(fn, destfile = "met_all.gz")
met <- data.table::fread("met_all.gz")
```

```{r, echo=T, message=FALSE, eval=FALSE}
download.file(
  "https://raw.githubusercontent.com/JSC370/jsc370-2022/main/labs/lab03/met_all.gz",
  destfile = "met_all.gz",
  method   = "curl",
  timeout  = 60
  )

met <- data.table::fread("met_all.gz")
```

### 2. Prepare the data

-   Remove temperatures less than -17C
-   Make sure there are no missing data in the key variables coded as 9999, 999, etc
-   Generate a date variable using the functions `as.Date()` (hint: You will need the following to create a date `paste(year, month, day, sep = "-")`).
-   Using the `data.table::week` function, keep the observations of the first week of the month.
-   Compute the mean by station of the variables `temp`, `rh`, `wind.sp`, `vis.dist`, `dew.point`, `lat`,`lon`, and `elev`.
-   Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
-   Create a categorical variable for elevation (breakpoint 252 m)

```{r, eval = TRUE, echo = FALSE}
# Remove temperatures less than -17C
met %>% filter(temp>=-17) %>%
# Make sure there are no missing data in the key variables coded as 9999, 999, etc
  na_if(9999) %>%
#  Generate a date variable using the functions `as.Date()` (hint: You will need the following to create a date `paste(year, month, day, sep = "-")`).
  mutate(date=as.Date(paste(year, month, day, sep="-"))) %>%
# Using the `data.table::week` function, keep the observations of the first week of the month.
  mutate(week=data.table::week(date)) %>% 
  filter(week==min(week)) %>%
# Compute the mean by station of the variables `temp`, `rh`, `wind.sp`, `vis.dist`, `dew.point`, `lat`,`lon`, and `elev`.
  group_by(USAFID) %>% 
  summarise(temp=mean(temp, na.rm = T),
            rh=mean(rh, na.rm = T),
            wind.sp=mean(wind.sp, na.rm = T),
            vis.dist=mean(vis.dist, na.rm = T),
            dew.point=mean(dew.point, na.rm = T),
            lat=mean(lat), lon=mean(lon),
            elev=mean(elev, na.rm = T)) %>%
# Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
  mutate(region = ifelse(lat > 39.71, "north", "south")) %>%
  mutate(region = paste(region, ifelse(lon > -98, "east", "west"))) %>%
# Create a categorical variable for elevation (breakpoint 252 m)
  mutate(elev_cat=ifelse(elev>252, "high", "low")) -> met_avg
```

### 3. Use `geom_violin` to examine the wind speed and dew point temperature by region

You saw how to use `geom_boxplot` in class. Try using `geom_violin` instead (take a look at the help). (hint: You will need to set the `x` aesthetic to 1)

-   Use facets
-   Make sure to deal with `NA` category
-   Describe what you observe in the graph

```{r, eval = TRUE, echo = FALSE}
met_avg %>% filter(! (region %in% NA)) %>% ggplot() + aes(x=1, y=wind.sp, fill=region) + geom_violin() + facet_wrap(~region)

met_avg %>% filter(! (region %in% NA)) %>% ggplot() + aes(x=1, y=dew.point, fill=region) + geom_violin() + facet_wrap(~region)
```
- I observed that the mean of wind speed in north west seems to be faster than north east, while the dew point temperature in north west seem to be higher than north east. 

### 4. Use `geom_jitter` with `stat_smooth` to examine the association between dew point temperature and wind speed by region

-   Color points by region
-   Make sure to deal with `NA` category
-   Fit a linear regression line by region
-   Describe what you observe in the graph

```{r, eval = TRUE, echo = FALSE}
met_avg %>% filter(! (region %in% NA)) %>% ggplot() + aes(x=dew.point, y=wind.sp, color=region) + geom_jitter() + 
# fit linear regression
stat_smooth(method=lm)
```
- I observed that the relationship between wind speed and dew point temp differs when regions are different.

### 5. Use `geom_bar` to create barplots of the weather stations by elevation category coloured by region

-   Bars by elevation category using `position="dodge"`
-   Change colors from the default. Color by region using `scale_fill_brewer` see [this](http://rstudio-pubs-static.s3.amazonaws.com/5312_98fc1aba2d5740dd849a5ab797cc2c8d.html)
-   Create nice labels on axes and add a title
-   Try a different theme
-   Describe what you observe in the graph
-   Make sure to deal with `NA`

```{r warning=FALSE, message=FALSE, eval = TRUE, echo = FALSE}
met_avg %>% filter(! (region %in% NA)) %>% ggplot() + aes(x=elev_cat, fill=region) + geom_bar(position = "dodge") + scale_fill_brewer(palette = "PuOr") + theme_minimal() + labs(title="Num of weather stations by elevation category and region", x="Elevation Category", y="count")
```
- I observed that most weather stations are built in low elevation in south west.

### 6. Use `stat_summary` to examine mean dew point and wind speed by region with standard deviation error bars

-   Make sure to remove `NA`
-   Use fun.data="mean_sdl" in `stat_summary`
-   Add another layer of `stats_summary` but change the geom to `"errorbar"` (see the help).
-   Describe the graph and what you observe

```{r, eval = TRUE, echo = TRUE}
met_avg %>% filter(! (region %in% NA)) %>% ggplot() + aes(x=region, y=dew.point) + stat_summary(fun.data = "mean_sdl", geom = "errorbar") + stat_summary(fun.data = "mean_sdl")

met_avg %>% filter(! (region %in% NA)) %>% ggplot() + aes(x=region, y=wind.sp) + stat_summary(fun.data = "mean_sdl", geom = "errorbar") + stat_summary(fun.data = "mean_sdl")
```

-   The mean Dew point temperature is about 16 in north east, 12.5 in north west, 21 in south east, and 13 in south west.
-   The mean Wind speed is about 1.7 in north east, 2.8 in north west, 1.7 in south east, and 3.2 in south west.

### 7. Make a map showing the spatial trend in relative h in the US

-   Make sure to remove `NA`
-   Use leaflet()
-   Make a colour palette with custom colours
-   Use `addMarkers` to include the top 10 places in relative h (hint: this will be useful `rank(-rh) <= 10`)
-   Add a legend

```{r, eval = TRUE, echo = FALSE}
top10 <- met_avg %>% arrange(desc(rh)) %>% slice_head(n=10)
rh_pal <- colorNumeric(c('blue', 'yellow', 'red'), domain = met_avg$rh, na.color = NA)
met_avg %>% drop_na() %>% leaflet() %>% addProviderTiles('OpenStreetMap') %>% addCircles(lat = ~lat, lng = ~lon, color=~rh_pal(rh), label=~paste(round(rh,2), 'rh'), opacity = 1, fillOpacity = 1, radius=~elev*10) %>% addMarkers(lat = ~lat, lng = ~lon, label = ~paste(round(rh,2), 'rh'), data = top10) %>% addLegend('bottomleft', pal = rh_pal, values=met_avg$rh, title="relative humidity", opacity=1)
```

-   Describe trend in RH across the US. The humidity is higher in east, and lower when closer to west. It gets very low around Nevada and Las Vegas.

### 8. Use a ggplot extension

-   Pick and extension (except cowplot) from [here](https://exts.ggplot2.tidyverse.org/gallery/) and make a plot of your choice using the met data (or met_avg)
-   Might want to try examples that come with the extension first (e.g. ggtech, gganimate, ggforce)
```{r}
library(ggtech)

# download.file("https://www.etsy.com/assets/type/Guardian-EgypTT-Text-Regular.ttf", "C:/Windows/Fonts/Guardian-EgypTT-Text-Regular.ttf", method="curl")

# extrafont::font_import(pattern = 'Guardian-EgypTT-Text-Regular.ttf', prompt=FALSE)

d <- qplot(temp, data = met_avg[met_avg$region %in%LETTERS[4:7], ], geom = "histogram", bins=30, fill = region)

d + theme_tech(theme="etsy") + 
  scale_fill_tech(theme="etsy") + 
  labs(title="Temperature")
```
- The required font can't be installed into the Win10 Fonts directory because the Font folder is a system folder under the protection. Therefore, the plot can't be shown without that font. I'll do a ggthemes library instead.
```{r}
library(ggthemes)

p <- met_avg %>% ggplot(aes(x=wind.sp, y=temp)) + geom_point() + stat_smooth(method=lm, color="red") + theme_economist() + scale_colour_economist() + ggtitle('temperature and wind speed are positively correlated')
p
```

